---
title: "Meta-analysis: Split-half reliability of the Serial Reaction Time task"
author: "Cátia Margarida Oliveira^[OSF page: https://osf.io/dpmfc/], Marianna E. Hayiou-Thomas, Lisa Henderson; University of York"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

<br> <br> This file is licensed under the terms and conditions of the CC-BY 4.0 INTERNATIONAL\
for details on the license, see\
<http://creativecommons.org/licenses/by/4.0/>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Import functions

```{r}
source("utils.R")
```

# Import libraries

```{r echo = T,warning=FALSE,message=FALSE}

listOfPackages <- c("rmarkdown","clubSandwich", "metafor","dplyr", "here", "apaTables", "tidyverse", "readr", "gridExtra")
ipak(listOfPackages)
```

# Data analysis

```{r echo = T}
# Read in file
Data <- read.csv("split-half reliability.csv")

# Check if including Buffington et al affects the findings

#Data <- subset(Data, Study != "Buffington et al. (2021)")

# Descriptives
Desc <- Data %>% group_by(Study, exp) %>%
  summarise(N = max(N), Age = mean(Age))
Desc

paste("Number of participants:", sum(Desc$N), "; Mean age (SD)",round(mean(Desc$Age),2), "(", round(sd(Desc$Age),2),") years;", "Range:", range(Desc$Age)[1],",", range(Desc$Age)[2])
```

# Preprocessing

```{r echo = TRUE}
#transform Spearman-Brown to Pearson's r
Data$cor <- ifelse(Data$Correction == "Spearman-Brown", Spearman_Brown_to_Pearson(Data$Split.half), Data$Split.half)

#transform correlations into Fisher Z and compute variance for effect sizes
Data  <- escalc(measure = "ZCOR", ri = cor, ni = N, data = Data, slab=paste(Authors, Year, sep=", "))

#Calculate standard error - Borenstein et al. (2009) 
Data$sei <- sqrt(Data$vi) 

# Preprocessing
Data[,c("Session", "exp")] <- lapply(Data[,c("Session", "exp")], factor)
Data[,c("Trials", "Age")] <- lapply(Data[,c("Trials", "Age")], scale)
Data$obs <- sequence(rle(Data$Study)$lengths)

#subset data - include only test-retest reliability for difference scores from Oliveira et al.
Data.DS <- Data[!(Data$Measure == "Regression slopes" & Data$Authors == "Oliveira et al."),] 

# within-study averaging assuming independent sampling errors
agg <- aggregate(Data, cluster=Data$exp, struct="ID")

```

# Data analysis

## Aggregated

```{r}
# fit random-effects model ignoring dependency
res.agg <- rma(yi, vi, data= agg)
res.agg

#reporter(res.agg)
```

## Multivariate model with covariance matrix - Only one measure per session/per study

This model includes only one measure per study, with difference scores being privileged when multiple indices of procedural learning were used as they are the most commonly used measure of procedural learning

```{r, warning=FALSE}
# construct approx V matrix assuming r=0.6 for the sampling errors
V.DS <- impute_covariance_matrix(Data.DS$vi, cluster=Data.DS$exp, r=0.6)

# fit multilevel model
res.ml.DS <- rma.mv(yi, V.DS, random = ~ 1 | Study/obs, data= Data.DS)
res.ml.DS

# fit multilevel model with nested obs within exp and exp within study
res2.DS <- rma.mv(yi, V.DS, random = ~ 1 | Study/exp/obs, data=Data.DS)
res2.DS

anova(res.ml.DS, res2.DS) #full shows sig better fit

#check fit
par(mfrow=c(2,2))
profile(res2.DS)

# compute average odds ratio (and corresponding 95% CI/PI)
predict(res2.DS, trans = transf.ztor, digits=2)

# RVE standard errors - guards agains model misspecification 
coef_test(res2.DS, vcov="CR2", cluster= Data.DS$Study)
robust(res2.DS, cluster=Data.DS$Study, clubSandwich=TRUE)
```

### Sensitivity analysis

```{r}

#combine effect sizes with different correlation sizes

compute_sensitivity.DS <- function(cor, vi, Study) {
  # construct approx V matrix assuming r=0.6 for the sampling errors
  V.DS <- impute_covariance_matrix(Data.DS$vi, cluster=Data.DS$exp, r=cor)

  # fit multivariate model
  res.DS.sen <- rma.mv(yi, V.DS, random = ~ 1 | Study/exp/obs, data=Data.DS)
  res.DS.sen

}

results.DS = list()
for (cor in  seq(.1, 1,.2))  {
    results.DS[[as.character(cor)]] = compute_sensitivity.DS(cor, vi, Study)
  }

print(results.DS)

```

## Multivariate model with covariance matrix - Only one measure per session/per study - without Oliveira et al.

We removed the studies conducted by Oliveira et al. because they contributed a lot of entries to the meta-analysis, thus by removing their results, we were able to assess whether their findings were influencing the overall estimates

```{r, warning=FALSE}

Data.DS.red <- Data.DS[Data.DS$Authors != "Oliveira et al.",]

# construct approx V matrix assuming r=0.6 for the sampling errors
V.DS.red <- impute_covariance_matrix(Data.DS.red$vi, cluster=Data.DS.red$exp, r=0.6)

# fit multilevel model with nested obs within exp and exp within study
res2.DS.red <- rma.mv(yi, V.DS.red, random = ~ 1 | Study/exp/obs, data=Data.DS.red)
res2.DS.red

#check fit
par(mfrow=c(2,2))
profile(res2.DS.red)

# compute average odds ratio (and corresponding 95% CI/PI)
predict(res2.DS.red, trans = transf.ztor, digits=2)

# RVE standard errors - guards agains model misspecification 
coef_test(res2.DS.red, vcov="CR2", cluster= Data.DS.red$Study)
```


## Multivariate model with covariance matrix - All data

By including the covariance matrix we are accounting for the fact that multiple entries per participant are being analysed, and thus these multiple entries are not independent from each other.

```{r, warning=FALSE}
# construct approx V matrix assuming r=0.6 for the sampling errors
V <- impute_covariance_matrix(Data$vi, cluster=Data$exp, r=0.6)

# fit multilevel model
res.ml <- rma.mv(yi, V, random = ~ 1 | Study/obs, data= Data)
res.ml

# fit multilevel model with nested obs within exp and exp within study
res2 <- rma.mv(yi, V, random = ~ 1 | Study/exp/obs, data=Data)
res2

anova(res.ml, res2) #no significant difference in fit

#check fit
par(mfrow=c(2,2))
profile(res2)

# compute average odds ratio (and corresponding 95% CI/PI)
predict(res2, trans = transf.ztor, digits=2)

# RVE standard errors - guards agains model misspecification 
coef_test(res2, vcov="CR2", cluster= Data$Study)

```

## Multivariate model with covariance matrix - Without Oliveira et al.

```{r, warning=FALSE}

red.ml <- Data[Data$Authors != "Oliveira et al.",]

# construct approx V matrix assuming r=0.6 for the sampling errors
V.red <- impute_covariance_matrix(red.ml$vi, cluster=red.ml$exp, r=0.6)

# fit multilevel model with nested obs within exp and exp within study
res2.red <- rma.mv(yi, V.red, random = ~ 1 | Study/exp/obs, data=red.ml)
res2.red

# compute average odds ratio (and corresponding 95% CI/PI)
predict(res2.red, trans = transf.ztor, digits=2)

# RVE standard errors - guards against model misspecification 
coef_test(res2.red, vcov="CR2", cluster= red.ml$Study)

```

### Forest plot

```{r}

png("forest_split-half-all.png", width = 1300, height = 1000)
pdf("forest_split-half-all.pdf", width = 19, height = 14)

# draw forest plot
forest(res2, header=TRUE, atransf=transf.ztor, refline=coef(res2), ilab.xpos=c(-10,-.5), cex=1.40, xlim = c(-1.3, 3))

### add text with Q-value, dfs, p-value, and I^2 statistic
text(-1.05, -1, pos=4, cex=1.40, bquote(paste("(Q = ",
   .(formatC(res2$QE, digits=2, format="f")), ", df = ", .(res2$k - res2$p),
   ", p <.001)")))
dev.off()
```

### Sensitivity analysis

```{r}

#combine effect sizes with different correlation sizes

compute_sensitivity <- function(cor, vi, Study) {
  # construct approx V matrix assuming r=0.6 for the sampling errors
  V <- impute_covariance_matrix(Data$vi, cluster=Data$exp, r=cor)

  # fit multivariate model
  res.sen <- rma.mv(yi, V, random = ~ 1 | Study/exp/obs, data= Data)
  res.sen

}

results = list()
for (cor in  seq(.1, 1,.2))  {
    results[[as.character(cor)]] = compute_sensitivity(cor, vi, Study)
  }

print(results)

```

## Meta regression {.tabset}

### Measure

```{r, warning=FALSE}

Data %>%
  group_by(Measure) %>%
  summarise(
    studies = n_distinct(Study),
    exp = n_distinct(exp),
    effects = n()
  )

Data$Measure <- factor(Data$Measure)

res.Measure <- rma.mv(yi, V, mods= ~ Measure -1, random = ~ 1 | Study/exp/obs, data= Data)
res.Measure

# guards against model misspecification
coef_test(res.Measure, vcov="CR2", cluster= Data$Study)

# Robust F-test - comparing whether the all levels are significantly different from zero
Wald_test(res.Measure, constraints = constrain_zero(1:2), vcov = "CR2", tidy = TRUE)

# Robust F-test - not enough datapoints for Ratio
Wald_test(res.Measure, constraints = constrain_pairwise(1:2), vcov = "CR2", tidy = TRUE)
```

### Age

```{r, warning=FALSE}

res.age <- rma.mv(yi, V, mods= ~ Age, random = ~ 1 | Study/exp/obs, data= Data)
res.age

# guards against model misspecification
coef_test(res.age, vcov="CR2", cluster= Data$Study)
```

### Trials

```{r, warning=FALSE}

res.trials <- rma.mv(yi, V, mods= ~ Trials, random = ~ 1 | Study/exp/obs, data= Data)
res.trials

# guards against model misspecification
coef_test(res.trials, vcov="CR2", cluster= Data$Study)
```

### SRTT type

```{r, warning=FALSE}

Data %>%
  group_by(SRTT.type) %>%
  summarise(
    studies = n_distinct(Study),
    exp = n_distinct(exp),
    effects = n())

res.SRTT.type <- rma.mv(yi, V, mods= ~ as.factor(SRTT.type)-1, random = ~ 1 | Study/exp/obs, data= Data)
res.SRTT.type

# guards against model misspecification
coef_test(res.SRTT.type, vcov="CR2", cluster= Data$Study)

# compares the levels against zero
Wald_test(res.SRTT.type, constraints = constrain_zero(1:3), vcov = "CR2")

# compares each level against each other
Wald_test(res.SRTT.type, constraints = constrain_pairwise(1:3), vcov = "CR2")
```

### ISI

```{r, warning=FALSE}

Data %>%
  group_by(ISI) %>%
  summarise(
    studies = n_distinct(Study),
    exp = n_distinct(exp),
    effects = n())

Data.ISI <- Data[Data$ISI != "500",]

# construct approx V matrix assuming r=0.6 for the sampling errors
V.ISI <- impute_covariance_matrix(Data.ISI$vi, cluster=Data.ISI$exp, r=0.6)

res.ISI <- rma.mv(yi, V.ISI, mods= ~ as.factor(ISI)-1, random = ~ 1 | Study/exp/obs, data= Data.ISI)
res.ISI

# guards against model misspecification
coef_test(res.ISI, vcov="CR2", cluster= Data.ISI$Study)

Wald_test(res.ISI, constraints = constrain_pairwise(1:2), vcov = "CR2")
```

# Publication bias and Influential/Outliers

## Aggregated effect sizes {.tabset}

### Funnel Plots

```{r, warning=FALSE}

### set up 2x2 array for plotting
par(mfrow=c(2,2))

### draw funnel plots
funnel(res.agg, main="Standard Error")
funnel(res.agg, yaxis="vi", main="Sampling Variance")
funnel(res.agg, yaxis="seinv", main="Inverse Standard Error")
funnel(res.agg, yaxis="vinv", main="Inverse Sampling Variance")
dev.off()

### funnel plot
funnel(res.agg, xlab = "Correlation coefficient")

### create contour enhanced funnel plot (with funnel centered at 0)
funnel(res.agg, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), refline=0, legend=TRUE)
```

### Egger's regression test/ PET-PEESE procedure

```{r, warning=FALSE}

# regression test for funnel plot asymmetry
regtest(res.agg) # uses the SEs as predictor by default
regtest(res.agg, predictor="vi") # use the variances as predictor
```

### Rank correlation test

```{r, warning=FALSE}

#Tests for bias
ranktest(res.agg, exact = FALSE)
```

### Outlier/influential points detection

```{r echo = T, out.width="50%", warning=FALSE}
par("mar")

### calculate influence diagnostics
inf <- influence(res.agg)

# save influential outlier case numbers in vector
which(inf$is.infl == "TRUE")
 
### plot the influence diagnostics
plot(inf)
```

## All effect sizes {.tabset}

### Funnel Plots

```{r, warning=FALSE}

### set up 2x2 array for plotting
par(mfrow=c(2,2))

### draw funnel plots
funnel(res2, main="Standard Error")
funnel(res2, yaxis="vi", main="Sampling Variance")
funnel(res2, yaxis="seinv", main="Inverse Standard Error")
funnel(res2, yaxis="vinv", main="Inverse Sampling Variance")
dev.off()

```

```{r contour li}

pdf("funnel.split-half.pdf", width = 14, height = 7)
par(mfrow=c(1,2))

### funnel plot
fp1 <- funnel(res2, back = "white", cex=1)

### create contour enhanced funnel plot (with funnel centered at 0)
funnel(res2, level=c(90, 95, 99), shade=c("white", "orange", "red"), refline=0, legend = TRUE,  back = "white") 

dev.off()
```

### Egger's regression test

```{r, warning=FALSE}
# run the regression test (with Vis as predictor) manually

#Reduced model
res.PET <- rma.mv(yi, vi, mods = ~ sei, random = ~ 1 | Study/exp, data=Data.DS)
res.PET

#Full model
res.ml.PET <- rma.mv(yi, V, mods = ~ sei, random = ~ 1 | Study/exp/obs, data=Data)
res.ml.PET

```

### Influential points {.tabset}

#### Effect size level

##### All data

```{r}
CD2 <- cooks.distance(res2)
plot(CD2, type="o", pch=19, xlab="Observed Outcome", ylab="Cook's Distance")
print(CD2)

DFbetas2 <- dfbetas(res2)
DFbetas2$scale <- scale(DFbetas2$intrcpt)
plot(DFbetas2$intrcpt, type="o", pch=19, xlab="Observed Outcome", ylab="DF betas")

```


#### Experiment level

##### DS data

```{r}
CD.cluster1 <- cooks.distance(res2.DS, cluster = Data.DS$exp)
plot(CD.cluster1, type="o", pch=19, xlab="Observed Outcome", ylab="Cook's Distance")
axis(side=1, at=seq_along(CD.cluster1), labels=as.numeric(names(CD.cluster1)))
print(names(CD.cluster1))

DFbetas.exp1 <- dfbetas(res2.DS, cluster = Data.DS$exp)
DFbetas.exp1$scale <- scale(DFbetas.exp1$intrcpt)
plot(DFbetas.exp1$intrcpt, type="o", pch=19, xlab="Observed Outcome", ylab="DF betas")
```

##### All data

```{r}
CD.cluster2 <- cooks.distance(res2, cluster = Data$exp)
plot(CD.cluster2, type="o", pch=19, xlab="Observed Outcome", ylab="Cook's Distance")
axis(side=1, at=seq_along(CD.cluster2), labels=as.numeric(names(CD.cluster2)))
print(names(CD.cluster2))

DFbetas.exp2 <- dfbetas(res2, cluster = Data$exp)
DFbetas.exp2$scale <- scale(DFbetas.exp2$intrcpt)
plot(DFbetas.exp2$intrcpt, type="o", pch=19, xlab="Observed Outcome", ylab="DF betas")
```

#### Study level

##### DS data

```{r}
CD.cluster3 <- cooks.distance(res2.DS, cluster = Data.DS$Study)
plot(CD.cluster3, type="o", pch=19, xlab="Observed Outcome", ylab="Cook's Distance")
axis(side=1, at=seq_along(CD.cluster3), labels=as.numeric(names(CD.cluster3)))

print(names(CD.cluster3))

DFbetas.study3 <- dfbetas(res2.DS, cluster = Data.DS$Study)
DFbetas.study3$scale <- scale(DFbetas.study3$intrcpt)
plot(DFbetas.study3$intrcpt, type="o", pch=19, xlab="Observed Outcome", ylab="DF betas")
```

##### All data

```{r}
CD.cluster4 <- cooks.distance(res2, cluster = Data$Study)
plot(CD.cluster4, type="o", pch=19, xlab="Observed Outcome", ylab="Cook's Distance")
axis(side=1, at=seq_along(CD.cluster4), labels=as.numeric(names(CD.cluster4)))
print(names(CD.cluster4))

DFbetas.study4 <- dfbetas(res2, cluster = Data$Study)
DFbetas.study4$scale <- scale(DFbetas.study4$intrcpt)
plot(DFbetas.study4$intrcpt, type="o", pch=19, xlab="Observed Outcome", ylab="DF betas")
```

### Outliers

#### DS data

```{r}
sr <- rstudent(res2.DS)

plot(sr$z, type="o", pch=19, xlab="Observed Outcome", ylab="Studentised residuals") + abline(h = 1.96, lty = 2, col = "tomato4")

print(sr$slab)
```

#### All data

```{r}
sr2 <- rstudent(res2)

plot(sr2$z, type="o", pch=19, xlab="Observed Outcome", ylab="Studentised residuals") + abline(h = 1.96, lty = 2, col = "tomato4")

print(sr2$slab)
```


# References

Viechtbauer, W. (2021). An overview of the metafor package with some interesting applications. Annual Meeting of the Society for Research Synthesis Methodology, Online. <https://www.wvbauer.com/doku.php/presentations>

<https://wviechtb.github.io/metafor/reference/rma.mv.html>

Assink, M., & Wibbelink, C. J. M. (2016). Fitting three-level meta-analytic models in R: A step-by-step tutorial. The Quantitative Methods for Psychology, 12(3), 154--174. <https://doi.org/10.20982/tqmp.12.3.p154>
