---
title: "Meta-analysis: Test-retest reliability of the Serial Reaction Time task"
author: "Cátia Margarida Oliveira^[OSF page: https://osf.io/dpmfc/], Marianna E. Hayiou-Thomas, Lisa Henderson; University of York"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

<br> <br> This file is licensed under the terms and conditions of the CC-BY 4.0 INTERNATIONAL\
for details on the license, see\
<http://creativecommons.org/licenses/by/4.0/>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import functions

```{r}
source("utils.R")
```

# Import libraries

```{r echo = T,warning=FALSE,message=FALSE}
listOfPackages <- c("rmarkdown","clubSandwich", "metafor","dplyr", "here", "apaTables")
ipak(listOfPackages)
```

# Data analysis

```{r echo = T}
# Read in file
Data <- read.csv("reliability_data.csv")

#transform correlations into Fisher Z and compute variance for effect sizes
Data  <- escalc(measure = "ZCOR", ri = Test.retest, ni = N, data = Data, slab=paste(Authors, Year, sep=", "))

#Create column distinguishing variables that account for speed from those which do not
Data$speed <- ifelse(Data$Measure != "RT difference score", 1,0)
  
#Calculate standard error - Borenstein et al. (2009) 
Data$sei <- sqrt(Data$vi) 

#subset data - include only test-retest reliability for difference scores if available
Data.DS <- Data[!(Data$Authors=="Oliveira et al." & Data$Trials_included == 1 | Data$Study=="West et al. (2021)" & Data$Measure == "Ratio" | Data$Contrast == 2| Data$Measure == "Regression slopes"),]

# within-study averaging assuming independent sampling errors
agg <- aggregate(Data.DS, cluster=Data.DS$exp, struct="ID")
```

# Descriptives

```{r echo = TRUE}

Desc <- Data %>% group_by(exp) %>% summarise(N = max(N), Age = mean(Age), trials = mean(Trials))

Desc

write.csv(Desc, "Desc.stats.tr.csv")
#write.csv(Desc, "desc.MA.csv")

paste("Number of participants:", sum(Desc$N), "; Mean age (SD)",round(mean(Desc$Age),2), "(", round(sd(Desc$Age),2), ") years;", "Range:", range(Desc$Age)[1],",", range(Desc$Age)[2])
```

#Preprocessing

```{r}
Data$group <- ifelse(Data$Age<15, "children", "adult")
Data$Trials <- scale(Data$Trials)
Data$Age <- scale(Data$Age)
```

# Data analysis

## Aggregated

```{r}
# fit random-effects model ignoring dependency
res.agg <- rma(yi, vi, data= agg)
res.agg

#reporter(res.agg)
```

## Multivariate model - only difference scores

```{r, warning=FALSE}
res.ml0 <- rma.mv(yi, vi, random = ~ 1 | Study, data= Data.DS)
res.ml1 <- rma.mv(yi, vi, random = ~ 1 | Study/exp, data=Data.DS)
res.ml1

# Compare models
anova(res.ml0, res.ml1)

# compute average odds ratio (and corresponding 95% CI/PI)
predict(res.ml1, trans = transf.ztor, digits=2)

# RVE SEs and p-values
coef_test(res.ml1, vcov="CR2", cluster= Data.DS$Study)
```

## Multivariate model - only difference scores without Oliveira et al. studies

```{r, warning=FALSE}

red.df <- Data.DS[Data.DS$Authors != "Oliveira et al.",]

res.reduced <- rma.mv(yi, vi, random = ~ 1 | Study/exp, data=red.df)
res.reduced

# compute average odds ratio (and corresponding 95% CI/PI)
predict(res.reduced, trans = transf.ztor, digits=2)

# RVE SEs and p-values
coef_test(res.reduced, vcov="CR2", cluster= red.df$Study)
```

### Forest plot - DS data

```{r}

png("forest_retest-DS.png", width = 1300, height = 1000)

# draw forest plot
forest(res.ml1, header=TRUE, atransf=transf.ztor, at=transf.rtoz(seq(-.3, 0.9,by=0.1)), refline=coef(res.ml1), ilab.xpos=c(-10,-.5), cex=1.40, xlim = c(-1.3, 2))

### add text with Q-value, dfs, p-value, and I^2 statistic
text(-1.10, -1, pos=4, cex=1.40, bquote(paste("(Q = ",
   .(formatC(res.ml1$QE, digits=2, format="f")), ", df = ", .(res.ml1$k - res.ml1$p),
   ", p =.005)")))
dev.off()

```

## Multivariate model with covariance matrix - All data

```{r, warning=FALSE}
Data$obs <- sequence(rle(Data$Study)$lengths)

# construct approx V matrix assuming r=0.6 for the sampling errors
V <- impute_covariance_matrix(Data$vi, cluster=Data$exp, r=0.6)

# fit multilevel model
res.ml3 <- rma.mv(yi, V, random = ~ 1 | Study/obs, data= Data)
res.ml3

# fit multilevel model with nested obs within exp and exp within study
res2 <- rma.mv(yi, V, random = ~ 1 | Study/exp/obs, data=Data)
res2

anova(res.ml3, res2) #no significant difference in fit

#check fit
par(mfrow=c(2,2))
profile(res2)

# compute average odds ratio (and corresponding 95% CI/PI)
predict(res2, trans = transf.ztor, digits=2)

# RVE standard errors - guards against model misspecification 
coef_test(res2, vcov="CR2", cluster= Data$Study)

```

### Forest plot - all data

```{r}

png("forest_retest-all.png", width = 1300, height = 1000)
pdf("forest_retest-all.pdf", width = 19, height = 14)

# draw forest plot
forest(res2, header=TRUE, atransf=transf.ztor, at=transf.rtoz(seq(-.4, 0.9,by=0.1)), refline=coef(res2), ilab.xpos=c(-10,-.5), cex=1.40, xlim = c(-1.3, 2))

### add text with Q-value, dfs, p-value, and I^2 statistic
text(-1.10, -1, pos=4, cex=1.40, bquote(paste("(Q = ",
   .(formatC(res2$QE, digits=2, format="f")), ", df = ", .(res2$k - res2$p),
   ", p <.001)")))

dev.off()

```

## Multivariate model with covariance matrix - Without Oliveira et al.

```{r, warning=FALSE}

red.ml <- Data[Data$Authors != "Oliveira et al.",]

red.ml$obs <- sequence(rle(red.ml$Study)$lengths)

# construct approx V matrix assuming r=0.6 for the sampling errors
V.red <- impute_covariance_matrix(red.ml$vi, cluster=red.ml$exp, r=0.6)

# fit multilevel model with nested obs within exp and exp within study
res2.red <- rma.mv(yi, V.red, random = ~ 1 | Study/exp/obs, data=red.ml)
res2.red

# compute average odds ratio (and corresponding 95% CI/PI)
predict(res2.red, trans = transf.ztor, digits=2)

# RVE standard errors - guards against model misspecification 
coef_test(res2.red, vcov="CR2", cluster= red.ml$Study)

```

### Sensitivity analysis

```{r}

#combine effect sizes with different correlation sizes

compute_sensitivity <- function(cor, vi, Study) {
  # construct approx V matrix assuming r=0.6 for the sampling errors
  V <- impute_covariance_matrix(Data$vi, cluster=Data$exp, r=cor)
  
  # fit multivariate model
  res.sen <- rma.mv(yi, V, random = ~ 1 | Study/exp/obs, data= Data)
  res.sen

}

results = list()
for (cor in  seq(.1, 1,.2))  {
    results[[as.character(cor)]] = compute_sensitivity(cor, vi, Study)
  }

print(results)

```

## Meta regression {.tabset}

### Measure

```{r, warning=FALSE}

Data %>%
  group_by(speed) %>%
  summarise(
    studies = n_distinct(Study),
    exp = n_distinct(exp),
    effects = n())

res.speed <- rma.mv(yi, V, mods= ~ as.factor(speed) -1, random = ~ 1 | Study/exp/obs, data= Data)
res.speed

# guards against model misspecification
coef_test(res.speed, vcov="CR2", cluster= Data$Study)

# Robust F-test
Wald_test(res.speed, constraints = constrain_pairwise(1:2), vcov = "CR2")
```

### Age

```{r, warning=FALSE}

res.age <- rma.mv(yi, V, mods= ~ Age, random = ~ 1 | Study/exp/obs, data= Data)
res.age

# guards against model misspecification
coef_test(res.age, vcov="CR2", cluster= Data$Study)

```

### Dichotomisation of age - Group

```{r echo = FALSE}

Data %>%
  group_by(group) %>%
  summarise(
    studies = n_distinct(Study),
    exp = n_distinct(exp),
    effects = n())

res.group <- rma.mv(yi, V, mods= ~ as.factor(group)-1, random = ~ 1 | Study/exp/obs, data= Data)
res.group

#RVE estimates - guards against model misspecification
coef_test(res.group, vcov="CR2", cluster= Data$Study)

# Robust F-test
Wald_test(res.group, constraints = constrain_pairwise(1:2), vcov = "CR2")
```

### Trials

```{r, warning=FALSE}

res.trials <- rma.mv(yi, V, mods= ~ Trials, random = ~ 1 | Study/exp/obs, data= Data)
res.trials

# guards against model misspecification
coef_test(res.trials, vcov="CR2", cluster= Data$Study)
```

### Sequence type

```{r, warning=FALSE}

Data %>%
  group_by(SRTT.type) %>%
  summarise(
    studies = n_distinct(Study),
    exp = n_distinct(exp),
    effects = n())

Data.SRTT.type <- subset(Data, SRTT.type != "alternating")

V.SRTT.type <- impute_covariance_matrix(Data.SRTT.type$vi, cluster=Data.SRTT.type$Study, r=cor)

res.SRTT.type <- rma.mv(yi, V.SRTT.type, mods= ~ as.factor(SRTT.type)-1, random = ~ 1 | Study/exp/obs, data = Data.SRTT.type)
res.SRTT.type

# guards against model misspecification
coef_test(res.SRTT.type, vcov="CR2", cluster= Data.SRTT.type$Study)

# Robust F-test
Wald_test(res.SRTT.type, constraints = constrain_pairwise(1:2), vcov = "CR2")
```

### ISI

```{r, warning=FALSE}

Data %>%
  group_by(ISI) %>%
  summarise(
    studies = n_distinct(Study),
    exp = n_distinct(exp),
    effects = n())

Data.ISI <- Data[Data$ISI == 0| Data$ISI == 250,]

V.ISI <- impute_covariance_matrix(Data.ISI$vi, cluster=Data.ISI$exp, r=cor)

res.ISI <- rma.mv(yi, V.ISI, mods= ~ as.factor(ISI)-1, random = ~ 1 | Study/exp/obs, data= Data.ISI)
res.ISI

# guards against model misspecification
coef_test(res.ISI, vcov="CR2", cluster= Data.ISI$Study)

# Robust F-test
Wald_test(res.ISI, constraints = constrain_pairwise(1:2), vcov = "CR2")
```

### Interval

```{r echo = FALSE}

Data.int <- Data[!is.na(as.numeric(Data$Interval)), ]
V.int <- impute_covariance_matrix(Data.int$vi, cluster=Data.int$exp, r=cor)

res.interval <- rma.mv(yi, V.int, mods= ~ scale(as.numeric(Interval)), random = ~ 1 | Study/exp/obs, data= Data.int)
res.interval

#RVE estimates - guards against model misspecification
coef_test(res.interval, vcov="CR2", cluster= Data.int$Study)
```

### Trials included

```{r echo = FALSE}

Data %>%
  group_by(Trials_included) %>%
  summarise(
    studies = n_distinct(Study),
    exp = n_distinct(exp),
    effects = n())

res.trials_included <- rma.mv(yi, V, mods= ~ as.factor(Trials_included)-1, random = ~ 1 | Study/exp/obs, data= Data)
res.trials_included

#RVE estimates - guards against model misspecification
coef_test(res.trials_included, vcov="CR2", cluster= Data$Study)

# Robust F-test
Wald_test(res.trials_included, constraints = constrain_pairwise(1:2), vcov = "CR2")
```

### Design

```{r echo = FALSE}

Data %>%
  group_by(design) %>%
  summarise(
    studies = n_distinct(Study),
    exp = n_distinct(exp),
    effects = n())

res.design <- rma.mv(yi, V, mods= ~ as.factor(design)-1, random = ~ 1 | Study/exp/obs, data= Data)
res.design

#RVE estimates - guards against model misspecification
coef_test(res.design, vcov="CR2", cluster= Data$Study)

# Robust F-test
Wald_test(res.design, constraints = constrain_pairwise(1:2), vcov = "CR2")
```


# Publication bias and Influential/Outliers

## Aggregated effect sizes {.tabset}

### Funnel Plots

```{r, warning=FALSE}

### set up 2x2 array for plotting
par(mfrow=c(2,2))

### draw funnel plots
funnel(res.agg, main="Standard Error")
funnel(res.agg, yaxis="vi", main="Sampling Variance")
funnel(res.agg, yaxis="seinv", main="Inverse Standard Error")
funnel(res.agg, yaxis="vinv", main="Inverse Sampling Variance")
dev.off()

### funnel plot
funnel(res.agg, xlab = "Reliability coefficient")

### create contour enhanced funnel plot (with funnel centered at 0)
funnel(res.agg, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), refline=0, legend=TRUE)
```

### Egger's regression test/ PET-PEESE procedure

```{r, warning=FALSE}

# regression test for funnel plot asymmetry
regtest(res.agg) # uses the SEs as predictor by default
regtest(res.agg, predictor="vi") # use the variances as predictor
```

### Rank correlation test

```{r, warning=FALSE}

#Tests for bias
ranktest(res.agg, exact = FALSE)
```

### Outlier/influential points detection

```{r echo = T, out.width="50%", warning=FALSE}
par("mar")

### calculate influence diagnostics
inf <- influence(res.agg)

# save influential outlier case numbers in vector
which(inf$is.infl == "TRUE")
 
### plot the influence diagnostics
plot(inf)
```

## All effect sizes {.tabset}

### Funnel Plots

```{r, warning=FALSE}

### set up 2x2 array for plotting
par(mfrow=c(2,2))

### draw funnel plots
funnel(res.ml1, main="Standard Error")
funnel(res.ml1, yaxis="vi", main="Sampling Variance")
funnel(res.ml1, yaxis="seinv", main="Inverse Standard Error")
funnel(res.ml1, yaxis="vinv", main="Inverse Sampling Variance")
dev.off()

```

```{r contour li}

png("funnel.test-retest.png", width = 1300, height = 700)
pdf("funnel.test-retest.pdf", width = 14, height = 7)

par(mfrow=c(1,2))
### funnel plot
fp1 <- funnel(res2, back = "white")

### create contour enhanced funnel plot (with funnel centered at 0)
fp2<- funnel(res2, level=c(90, 95, 99), shade=c("white", "orange", "red"), refline=0, legend = TRUE, back = "white")

dev.off()
```

### Egger's regression test

```{r, warning=FALSE}

res.PET <- rma.mv(yi, vi, mods = ~ sei, random = ~ 1 | Study/exp, data=Data.DS)
res.PET

# run the regression test (with Vis as predictor) manually
res.ml.PET <- rma.mv(yi, V, mods = ~ sei, random = ~ 1 | Study/exp/obs, data=Data)
res.ml.PET

```

### Influential points {.tabset}

#### Effect size level

##### All data

```{r}
CD2 <- cooks.distance(res2)
plot(CD2, type="o", pch=19, xlab="Observed Outcome", ylab="Cook's Distance")
print(CD2)

DFbetas2 <- dfbetas(res2)
plot(DFbetas2$intrcpt, type="o", pch=19, xlab="Observed Outcome", ylab="DF betas")

DFbetas2$scale <- scale(DFbetas2$intrcpt)

```


#### Experiment level

##### DS data

```{r}
CD.cluster1 <- cooks.distance(res.ml1, cluster = Data.DS$exp)
plot(CD.cluster1, type="o", pch=19, xlab="Observed Outcome", ylab="Cook's Distance")
axis(side=1, at=seq_along(CD.cluster1), labels=as.numeric(names(CD.cluster1)))
print(names(CD.cluster1))

DFbetas.exp1 <- dfbetas(res.ml1, cluster = Data.DS$exp)
plot(DFbetas.exp1$intrcpt, type="o", pch=19, xlab="Observed Outcome", ylab="DF betas")
DFbetas.exp1$scale <- scale(DFbetas.exp1$intrcpt)

```

##### All data

```{r}
CD.cluster2 <- cooks.distance(res2, cluster = Data$exp)
plot(CD.cluster2, type="o", pch=19, xlab="Observed Outcome", ylab="Cook's Distance")
axis(side=1, at=seq_along(CD.cluster2), labels=as.numeric(names(CD.cluster2)))
print(names(CD.cluster2))

DFbetas.exp2 <- dfbetas(res2, cluster = Data$exp)
plot(DFbetas.exp2$intrcpt, type="o", pch=19, xlab="Observed Outcome", ylab="DF betas")
DFbetas.exp2$scale <- scale(DFbetas.exp2$intrcpt)
```

#### Study level

##### DS data
```{r}
CD.cluster3 <- cooks.distance(res.ml1, cluster = Data.DS$Study)
plot(CD.cluster3, type="o", pch=19, xlab="Observed Outcome", ylab="Cook's Distance")
axis(side=1, at=seq_along(CD.cluster3), labels=as.numeric(names(CD.cluster3)))

print(names(CD.cluster3))

DFbetas.study3 <- dfbetas(res.ml1, cluster = Data.DS$Study)
plot(DFbetas.study3$intrcpt, type="o", pch=19, xlab="Observed Outcome", ylab="DF betas")

DFbetas.study3$scale <- scale(DFbetas.study3$intrcpt)

```


##### All data
```{r}
CD.cluster4 <- cooks.distance(res2, cluster = Data$Study)
plot(CD.cluster4, type="o", pch=19, xlab="Observed Outcome", ylab="Cook's Distance")
axis(side=1, at=seq_along(CD.cluster4), labels=as.numeric(names(CD.cluster4)))
print(names(CD.cluster4))

DFbetas.study4 <- dfbetas(res2, cluster = Data$Study)
plot(DFbetas.study4$intrcpt, type="o", pch=19, xlab="Observed Outcome", ylab="DF betas")
DFbetas.study4$scale <- scale(DFbetas.study4$intrcpt)

```

### Outliers

#### DS data

```{r}
sr <- rstudent(res.ml1)

plot(sr$z, type="o", pch=19, xlab="Observed Outcome", ylab="Studentised residuals") +abline(h = 1.96, lty = 2, col = "tomato4")

print(sr$slab)
```

#### All data

```{r}
sr2 <- rstudent(res2)

plot(sr2$z, type="o", pch=19, xlab="Observed Outcome", ylab="Studentised residuals") +abline(h = 1.96, lty = 2, col = "tomato4")

print(sr2$slab)
```


# References

Viechtbauer, W. (2021). An overview of the metafor package with some interesting applications. Annual Meeting of the Society for Research Synthesis Methodology, Online. <https://www.wvbauer.com/doku.php/presentations>

<https://wviechtb.github.io/metafor/reference/rma.mv.html>

Assink, M., & Wibbelink, C. J. M. (2016). Fitting three-level meta-analytic models in R: A step-by-step tutorial. The Quantitative Methods for Psychology, 12(3), 154--174. <https://doi.org/10.20982/tqmp.12.3.p154>
